|

### 2024-01-28 rmds/figures.Rmd
changes

---
output: pdf_document
---

@@

```{r out.width = '100%', fig.cap = "Approach used to generate simulated data and test the performance of four methods for calculating maternity parameters. "}
knitr::include_graphics(here("img","figure1.png"))
```

\newpage
```{r fig.width = 7, fig.height= 5.5}
data <- read_rds(here("data", "simulations-pivot.Rds"))

ggplot() + 
  geom_point(data = data, aes(x = rel_error_c, y = rel_error_m50, col = repro_freq)) +
  scale_color_viridis_d(option = "D", direction = 1) + 
  facet_grid(species~method_long) + 
  ylim(-30, 50) +
  guides(colour = guide_legend(bquote(atop(paste(plain("True "),italic(P)[Max]),("Reproductive frequency"))))) + 
  theme_bw() +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center") + 
  xlab(expression("% relative error "~hat(italic(P)[Max]))) +
  ylab(expression("% relative error "~hat(italic(L^"'")[50])))
```

Figure 2.  Per cent relative error (bias) in parameter estimates for $\hat{L'_{50}}$ and $\hat{P_{Max}}$ for 3PLF methods. Each point represents parameter estimates one iteration of simulated data. Bias tended to increase as the true underlying $P_{Max}$ decreased. 
\newpage

```{r}
data <- read_rds(here("data", "simulations-slim.Rds"))

data <- data %>% 
    mutate(method_long = case_when(
    method %in% "2" ~ "2PLF - maturity",
    method %in% "2a" ~ "2PLF - maternity",
    method %in% "3" ~ "3PLF - free",
    method %in% "3a" ~ "3PLF - fixed"
  )) %>%
  mutate(mesh_name_long = case_when(
    mesh_name %in% "high" ~ "High selectivity",
    mesh_name %in% "low" ~ "Low selectivity"
  )) %>% 
  filter(mesh_name %in% "high", par_name %in% "r_0") %>% 
  filter(Nsamples %in% c(50, 250, 1000))

data <- data %>% 
  mutate(method_long = fct_relevel(
    method_long, "3PLF - free",
    "3PLF - fixed",
    "2PLF - maternity",
    "2PLF - maturity"
  ))

ggplot(data) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_boxplot(aes(x = Nsamples, y = abs_error, fill = repro_freq), outlier.shape = NA) +
  facet_grid(method_long ~ species) + 
  scale_color_brewer(palette = "Spectral") + 
  guides(fill = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
  cowplot::theme_half_open() +
  panel_border() +
  background_grid() +
  labs(x = "Sample size", y = "Median % absolute error") +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center") + 
  ylim(0, 80)
```

\newpage
```{r fig.width = 10, fig.height= 7}
# Load data - Sandbar
data <- read_rds(here("data", "empirical-plumbeus.Rds"))
preds <- read_rds(here("data", "empirical-predictions.Rds"))

# Bin data into length categories for plotting
brks = seq(40, 220, 8)
data_binned <- data %>%
  mutate(x_bin = findInterval(x, brks)) %>% 
  mutate(x_bin = (brks[x_bin] + brks[x_bin + 1]) / 2) %>% 
  group_by(x_bin) %>%
  summarise(p = sum(z)/ n())

preds1 <- filter(preds, species %in% "C. plumbeus", label %in% "3PLF - free") %>% 
  mutate(label2 = "italic('C. plumbeus ')~`- 3PLF  free`")

p1 <- ggplot()  + 
  geom_ribbon(data = preds1, aes(x = x, ymin = lower, ymax = upper), fill = "grey70") + 
  geom_line(data = preds1, aes(x = x, y = y)) +
  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
  geom_rug(data = filter(data, z == 1), aes(x = x), sides = "t") +
  geom_rug(data = filter(data, z == 0), aes(x = x), sides = "b") +
  ylim(0, 1) +
  theme_cowplot() + 
  labs(x = NULL, y = NULL) + 
  facet_wrap(~label2, labeller = label_parsed) + 
  theme(strip.text = element_text(hjust = 0.5))

preds2 <- filter(preds, species %in% "C. plumbeus", label %in% "3PLF - fixed (PMax = 0.5)") %>%
  mutate(label2 = "italic('C. plumbeus')~` - 3PLF fixed`")

# Plot
p2 <- ggplot()  + 
  geom_ribbon(data = preds2, aes(x = x, ymin = lower, ymax = upper), fill = "grey70") + 
  geom_line(data = preds2, aes(x = x, y = y)) +
  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
  geom_rug(data = filter(data, z == 1), aes(x = x), sides = "t") +
  geom_rug(data = filter(data, z == 0), aes(x = x), sides = "b") +
  ylim(0, 1) +
  theme_cowplot() + 
  labs(x = NULL, y = NULL) + 
  facet_wrap(~label2, labeller = label_parsed) + 
  theme(strip.text = element_text(hjust = 0.5))

# Load data - blacktip
data <- read_rds(here("data", "empirical-limbatus.Rds"))

# Bin data into length categories for plotting
brks = seq(40, 160, 8)
data_binned <- data %>%
  mutate(x_bin = findInterval(x, brks)) %>% 
  mutate(x_bin = (brks[x_bin] + brks[x_bin + 1]) / 2) %>% 
  group_by(x_bin) %>%
  summarise(p = sum(z)/ n())

preds3 <- filter(preds, species %in% "C. limbatus", label %in% "3PLF - free") %>%
  mutate(label2 = "italic('C. limbatus')~` - 3PLF free`")

# Plot
p3 <- ggplot()  + 
  geom_ribbon(data = preds3, aes(x = x, ymin = lower, ymax = upper), fill = "grey70") + 
  geom_line(data = preds3, aes(x = x, y = y)) +
  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
  geom_rug(data = filter(data, z == 1), aes(x = x), sides = "t") +
  geom_rug(data = filter(data, z == 0), aes(x = x), sides = "b") +
  ylim(0, 1) +
  theme_cowplot() + 
  labs(x = NULL, y = NULL) + 
  facet_wrap(~label2, labeller = label_parsed) + 
  theme(strip.text = element_text(hjust = 0.5))

preds4 <- filter(preds, species %in% "C. limbatus", label %in% "3PLF - fixed (PMax = 0.5)") %>%
  mutate(label2 = "italic('C. limbatus')~` - 3PLF fixed`")

p4 <- ggplot()  + 
  geom_ribbon(data = preds4, aes(x = x, ymin = lower, ymax = upper), fill = "grey70") + 
  geom_line(data = preds4, aes(x = x, y = y)) +
  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
  geom_rug(data = filter(data, z == 1), aes(x = x), sides = "t") +
  geom_rug(data = filter(data, z == 0), aes(x = x), sides = "b") +
  ylim(0, 1) +
  theme_cowplot() + 
  labs(x = NULL, y = NULL)  + 
  facet_wrap(~label2, labeller = label_parsed) + 
  theme(strip.text = element_text(hjust = 0.5))

grid.arrange(p3, p4, p1, p2, left = textGrob("Proportion in maternal condition", gp = gpar(fontsize = 16, font = 8), rot = 90), bottom = textGrob("Fork length (cm)", gp = gpar(fontsize = 16, font = 8)))
```
*Figure 3. Comparison of 3PLF-free and -fixed methods used to estimate maternal parameters for blacktip shark, *C. limbatus*, and sandbar shark, *C. plumbeus* in the Gulf of Mexico and Western North Atlantic. Sold line is the expected proportion in maternal condition at fork length. The grey shaded region denotes 95% confidence intervals based on bootstrap resampling. A fixed value of $P_{Max} = 0.5$ was used for both species.*

>>>>

---
output: 
  pdf_document:
    latex_engine: "pdflatex"
---

@@

```{r out.width = '100%'}
knitr::include_graphics(here("img","figure1.png"))
```

Figure 1. Approach used to generate simulated data and test the performance of four methods for calculating maternity parameters. Illustrations © R.Swainston/www.anima.net.au

```{r fig.width = 7, fig.height= 5.5}
data <- read_rds(here("data", "simulations-pivot.Rds"))

ggplot() + 
  geom_point(data = data, aes(x = rel_error_c, y = rel_error_m50, col = repro_freq)) +
  scale_color_viridis_d(option = "D", direction = 1) + 
  facet_grid(species~method_long) + 
  ylim(-30, 50) +
  guides(colour = guide_legend(bquote(atop(paste(plain("True "),italic(P)[Max]),("Reproductive frequency"))))) + 
  theme_bw() +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center") + 
  xlab(expression("% relative error "~hat(italic(P)[Max]))) +
  ylab(expression("% relative error "~hat(italic(L^"'")[50])))
```
Figure 2. Bias (per cent relative error) in parameter estimates for $\hat{L'_{50}}$ and $\hat{P_{Max}}$ for 3PLF methods. Each point represents parameter estimates from one iteration of simulated data. Bias tended to increase as the true underlying $P_{Max}$ decreased.
\
\
\newpage

```{r fig.width = 7, fig.height= 5.5}
data <- read_rds(here("data", "simulations-summarised.Rds"))

data %>% 
  filter(par_name %in% "c") %>% 
  filter(method_long %in% c("3PLF - estimated", "3PLF - fixed")) %>% 
  filter(mesh_name %in% "high") %>% 
  ggplot() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_point(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) +
  geom_line(aes(as.factor(Nsamples), abs_error, col = repro_freq, group = repro_freq)) + 
  facet_grid(species ~ method_long) + 
  scale_color_brewer(palette = "Spectral") + 
  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
  theme_bw() + 
  labs(x = "Sample size", y = "Mean % absolute relative error") +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
```
Figure 3. Precision (per cent absolute error) in parameter estimates  of $\hat{P_{Max}}$ for 3PLF methods. Large sample sizes were needed to accurately estimate $\hat{P_{Max}}$ and precision decreased as the duration of the reproductive cycle increased. Each point reflects a mean value from 300 simulated data sets. 
\
\
\newpage

```{r fig.width = 7, fig.height= 5.5}
data %>% 
  filter(par_name %in% "m50", method_long %in% c("3PLF - estimated", "3PLF - fixed")) %>% 
  filter(mesh_name %in% "high") %>% 
  ggplot() +
  geom_hline(yintercept = 50, linetype = "dashed") +
  geom_point(aes(as.factor(Nsamples), int_coverage, col = repro_freq, group = repro_freq)) +
  geom_line(aes(as.factor(Nsamples), int_coverage, col = repro_freq, group = repro_freq)) + 
  facet_grid(species ~ method_long, scales = "free_y") + 
  scale_color_brewer(palette = "Spectral") + 
  guides(colour = guide_legend(bquote(atop(italic(P)[Max],("Reproductive frequency"))))) + 
  theme_bw() + 
  labs(x = "Sample size", y = "% interval coverage") +
  theme(legend.position = "bottom", legend.direction = "horizontal", legend.justification = "center")
```
Figure 4. Confidence interval coverage for $\hat{L'_{50}}$ for 3PLF methods (high selectivity scenarios). Figure shows the percentage of simulations (n = 300) where the true parameter value fell within the 50% bootstrap confidence interval.
\
\
\newpage

```{r fig.width = 9, fig.height= 3.5}

read_rds(here("data", "simulations-summarised.Rds")) %>% 
  filter(!(method_long %in% "2PLF - maternity" & repro_freq %in% "1 (Annual)")) %>% 
  mutate(Nsamples_sel = fct_cross(as.factor(mesh_name), as.factor(Nsamples))) %>% 
  filter(par_name %in% "r_0") %>% group_by(species, Nsamples_sel, repro_freq) %>% 
  slice_min(abs(rel_error)) %>% 
  ggplot() + 
  geom_tile(aes(x = as.factor(Nsamples_sel), y = repro_freq, fill = method_long)) + 
  scale_fill_viridis_d() + 
  guides(fill = guide_legend(title = "Method")) + 
  facet_wrap(~species, nrow = 1) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) + 
  labs(x = "Selectivity: Sample size", y = bquote(atop(italic(P)[Max],("Reproductive frequency")))) 
```
Figure 5. Performance of alternative maternity functions in minimising bias in calculations of $R_0$. The preferred method was that which minimised bias, $|\text{relative error}|$ across 300 simulated datasets. Note 2PLF-maternity (Annual) scenarios were excluded for this comparison.
\
\
\newpage

```{r fig.width = 9, fig.height= 3.5}

read_rds(here("data", "simulations-summarised.Rds")) %>% 
  filter(!(method_long %in% "2PLF - maternity" & repro_freq %in% "1 (Annual)")) %>% 
  mutate(Nsamples_sel = fct_cross(as.factor(mesh_name), as.factor(Nsamples))) %>% 
  filter(par_name %in% "r_0") %>% group_by(species, Nsamples_sel, repro_freq) %>% 
  slice_min(abs_error) %>%
  ggplot() + 
  geom_tile(aes(x = as.factor(Nsamples_sel), y = repro_freq, fill = method_long)) + 
  scale_fill_viridis_d() + 
  guides(fill = guide_legend(title = "Method")) + 
  facet_wrap(~species, nrow = 1) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) + 
  labs(x = "Selectivity: Sample size", y = bquote(atop(italic(P)[Max],("Reproductive frequency")))) 
```
Figure 6. Performance of alternative maternity functions in accurately calculating $R_0$. The best performing method was that which minimised mean absolute error across 300 simulated datasets. Note 2PLF-maternity (Annual) scenarios were excluded for this comparison.
\
\newpage

```{r fig.width = 4.5, fig.height= 5}
# Load data - Sandbar
data <- read_rds(here("data", "empirical-plumbeus.Rds"))
preds <- read_rds(here("data", "empirical-predictions.Rds"))

# Bin data into length categories for plotting
brks = seq(40, 220, 7.5)
data_binned <- data %>%
  mutate(x_bin = findInterval(x, brks)) %>% 
  mutate(x_bin = (brks[x_bin] + brks[x_bin + 1]) / 2) %>% 
  group_by(x_bin) %>%
  summarise(p = sum(z)/ n())

preds1 <- filter(preds, species %in% "C. plumbeus", label %in% "3PLF - estimated") %>% 
  mutate(label2 = "3PLF - estimated")

p1 <- ggplot()  + 
  geom_ribbon(data = preds1, aes(x = x, ymin = lower, ymax = upper), fill = "grey70") + 
  geom_line(data = preds1, aes(x = x, y = y)) +
  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
  geom_rug(data = filter(data, z == 1), aes(x = x), sides = "t") +
  geom_rug(data = filter(data, z == 0), aes(x = x), sides = "b") +
  ylim(0, 1) +
  theme_bw() + 
  labs(x = NULL, y = NULL) + 
  facet_wrap(~label2) + 
  theme(strip.text = element_text(hjust = 0.5))

preds2 <- filter(preds, species %in% "C. plumbeus", label %in% "3PLF - fixed (PMax = 0.5)") %>%
  mutate(label2 = "3PLF - fixed")

# Plot
p2 <- ggplot()  + 
  geom_ribbon(data = preds2, aes(x = x, ymin = lower, ymax = upper), fill = "grey70") + 
  geom_line(data = preds2, aes(x = x, y = y)) +
  geom_point(data = data_binned, aes(x = x_bin, y = p), col = "black") + 
  geom_rug(data = filter(data, z == 1), aes(x = x), sides = "t") +
  geom_rug(data = filter(data, z == 0), aes(x = x), sides = "b") +
  ylim(0, 1) +
  theme_bw() + 
  labs(x = NULL, y = NULL) + 
  facet_wrap(~label2) + 
  theme(strip.text = element_text(hjust = 0.5))

grid.arrange(p1, p2, left = textGrob("Proportion in maternal condition", gp = gpar(fontsize = 12, font = 8), rot = 90), bottom = textGrob("Fork length (cm)", gp = gpar(fontsize = 12, font = 8)))
```

Figure 7. Comparison of 3PLF-estimated and 3PLF-fixed methods used to estimate maternal parameters for sandbar shark, *C. plumbeus*, in the Gulf of Mexico and Western North Atlantic. Solid line is the expected proportion in maternal condition at length, $\Psi'(L)$. The grey shaded region denotes 95% confidence intervals based on bootstrap resampling. $P_{Max}$ was fixed at 0.5 in the lower panel
